Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-06-26 propriety_section.Rmd
adding marginal likelihood section

Our models consist of several parameters. Therefore, computing the Bayes factor of, for example, the normal model versus the log-normal model becomes somewhat difficult. Let the vector of all parameters of the normal model be $\bm{\theta_\text{n}}$ and the vector of all parameters for the log-normal model be $\bm{\theta_{\text{l}}}$. With the prior distributions for the two models, $f(\bm{\theta_\text{n}})$ and $f(\bm{\theta_\text{l}})$, the Bayes factor can be calculated by the following ratio of marginal likelihoods:

\begin{equation}
\text{BF}_{\text{nl}} = \frac{P(\bm{Y} | \mathcal{M}_n)}{P(\bm{Y} | \mathcal{M}_{\text{l}})} = \frac{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{n}, \bm{\theta_{\text{n}}})f(\bm{\theta_{\text{n}}})d\bm{\theta_{\text{n}}}}{\int_{-\infty}^{\infty}P(\bm{Y} | \mathcal{M}_\text{l}, \bm{\theta_{\text{l}}})f(\bm{\theta_{\text{l}}})d\bm{\theta_{\text{l}}}}. 
\end{equation}

>>>>

# Model Comparison

The symbolic distance hypothesis postulates that response times (RTs) are slower when digits are closer to 5 (digit effects), and that RTs may be influenced by whether the target digit is smaller or greater than 5 (side effect). Solely using Bayesian estimation, we are not able to directly test these hypotheses. Credible intervals often contain zero, but they always also contain lots of other values that have a chance of being the "true" parameter value, making posterior-based intervals unsuitable to test hypotheses. Therefore, our goal for the following section is to set up a Bayesian hypothesis test for the parameters of interest in our model. Specifically, we will achieve this by comparing the predictive accuracy of a null model where we set one or some parameters to zero, and and effects models where these parameters are unconstrained. \Revision{This approach is called Bayes factor model comparison.}

## Bayes factor

\Revision{Rather than submerging us in technical details or providing a formal justification of the Bayesian approach, we provide an intuition of Bayesian model comparison. A formal justification of Bayes factors is provided by @jeffreys1961theory, @kass1995bayes and @rouder2018teaching. The Bayes factor (BF) is a measure of how well one model predicts the data compared to another.}

\Revision{Bayesian analysis allows for predictions of data from any model because models are fully specified with a probability distribution for the data and priors for all parameters. Let the vector of all parameters of a model be $\bm{\theta}$. With the prior distributions these parameters, $f(\bm{\theta})$, the prediction for data from the model can be computed by integrating over all parameters:}

\begin{equation}
p(\bm{Y} | \mathcal{M}) = \int_{\bm{\Theta}} p(\bm{Y} | \mathcal{M}, \bm{\theta})f(\bm{\theta})d\bm{\theta}. 
\end{equation}

\Revision{Note that $\int_{\bm{\Theta}}$ refers to a multidimensional integral over the entire parameter space. As our models consist of many parameters computing these predictions for any one model becomes somewhat difficult. Once the integral is computed, the resulting probability distribution of the data, $p(\bm{Y} | \mathcal{M})$ is not a function of the model parameters anymore. This marginal probability distribution is therefore also referred to as the marginal likelihood. Before observing the data, the marginal likelihood serves as prediction for possible data; once the data are observed, the function tells us how well these specific observations were predicted by the model. The Bayes factor then is the ratio of marginal likelihoods of two models, or the relative predictive accuracy of $\mathcal{M}_1$ over $\mathcal{M}_2$:}

\begin{equation}
\text{BF}_{1,2} = \frac{P(\bm{Y} | \mathcal{M}_1)}{P(\bm{Y} | \mathcal{M}_2)}. 
\end{equation}

\Revision{There is a second popular interpretation of the Bayes factor as the relative evidence of two competing models. One of the key insights from Bayes' rule is that these to things, the relative predictive accuracy of two models for the data, and the relative evidence from the data for the two models, are one and the same [@Rouder:Morey:2018].}

Bayes factors have a few very convenient characteristics. One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13. For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers and colleagues [-@wagenmakers2018bayesian].

\newpage
